---
title: "Anaerobic CUE - 16S"
subtitle: "02 Decontaminate dataset"
author: "Roey Angel"
email: "roey.angel@bc.cas.cz"
date: "`r Sys.Date()`"
bibliography: references.bib
link-citations: yes
csl: fems-microbiology-ecology.csl
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    keep_md: false
    number_sections: false
    highlight: "pygments"
    theme: "flatly"
    dev: "png"
    df_print: "kable"
    fig_caption: true
    code_folding: "show"
---

```{r libraries, include=F}
# Load libraries
#.libPaths(c('~/R/library', .libPaths())) # Uncomment if you have no write access to R path

repo <- "http://cran.wu.ac.at"
lib.loc <- Sys.getenv("R_LIBS_USER")

update.packages(
    lib.loc, 
    repos = repo,
    ask = FALSE
)

.cran_libs <- c(
  "remotes", # R Package Installation from Remote Repositories, Including 'GitHub'
  "knitr", # A General-Purpose Package for Dynamic Report Generation in R
  "kableExtra", # Construct Complex Table with 'kable' and Pipe Syntax
  "extrafont", # for extra figure fonts
  "rmarkdown", # Dynamic Documents for R
  "tidyverse", # for dplyr forcats ggplot2 readr tibble
  "cowplot", # wrappers for ggplot
  "magrittr", # pipes
  "scales", # Generic plot scaling methods
  "svglite" # for svg files
) 

.cran_inst <- .cran_libs %in% installed.packages()
if (any(!.cran_inst)) {
   install.packages(.cran_libs[!.cran_inst], 
                    repos = repo,
                    lib = lib.loc)
}

.bioc_libs <- c(
  "phyloseq" # Handling and analysis of high-throughput phylogenetic sequence data
)

if (!"phyloseq" %in% installed.packages()) {
      source(
        "https://raw.githubusercontent.com/joey711/phyloseq/master/inst/scripts/installer.R",
        local = TRUE
      )
}

.bioc_inst <- .bioc_libs %in% installed.packages()
if (any(!.bioc_inst)) {
   if (!requireNamespace("BiocManager", quietly = TRUE))
   install.packages("BiocManager")
   BiocManager::install(ask = F, lib = lib.loc)  # upgrade bioC packages
   BiocManager::install(.bioc_libs[!.bioc_inst], ask = F, lib = lib.loc)
}

.local_libs <- c(
  my.tools = "my.tools_0.4_R_x86_64-pc-linux-gnu.tar.gz"
)

.local_inst <- names(.local_libs) %in% installed.packages()
if (any(!.local_inst)) {
   install.packages(paste0("~/R/", .local_libs[!.inst]) ,repos = NULL, type = "source", lib = lib.loc)
}

.github_libs <- c(
  "benjjneb/decontam", # dentify Contaminants in Marker-gene and Metagenomics Sequencing Data
  "clauswilke/ggtext", # Improved text rendering for ggplot2.
  "mikemc/speedyseq" # Faster implementations of common phyloseq functions
)

.github_lib_names <- stringr::str_replace(.github_libs, ".*/(.*)$", "\\1")

.github_inst <- .github_lib_names %in% installed.packages()
if (any(!.github_inst)) {
  install_github(.github_libs[!.github_inst],
                           lib = lib.loc,
                           dependencies = TRUE)
}

# Load packages into session, and print package version
(loaded.libs <- sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), require, character.only = TRUE))
if (!all(loaded.libs)) {stop(paste("Package(s):", names(loaded.libs[loaded.libs == FALSE]), "could not be loaded"))}
sapply(c(.cran_libs, .bioc_libs, names(.local_libs), .github_lib_names), packageVersion)
```
```{r style settings, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
options(width = 90, knitr.table.format = "html") 
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  dev = "png",
  fig.ext = "png",
  dpi = 600,
#  fig.width = 12,
#  fig.height = 8,
  cache.path = "02_Decontamination_cache/",
  fig.path = "02_Decontamination_figures/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 14
font_import(pattern = "DejaVuSans", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name))
```

```{r functions, include=F}
```

[roey.angel@bc.cas.cz](mailto: roey.angel@bc.cas.cz) 

## Decontamination
Decontamination of sequence library based on [Introduction to decontam](https://benjjneb.github.io/decontam/vignettes/decontam_intro.html) and Davis and colleagues [-@davis_simple_2018]. Decontamination is based on correlating sequence abundance frequencies to initial DNA concentrations used for PCR and also on examining sequence abundance prevalence patterns in the negative control samples.

### Setting general parameters:
```{r general parameters}
set.seed(1000)
metadata_path <- "./"
data_path <- "./DADA2_pseudo/"
Metadata_table <- "Metadata.csv"
Seq_table <- "DADA2.seqtab_nochim.tsv"
# Metadata_table <- ""
```

### Reading in raw data and generate phyloseq object
```{r load data, cache=T}
# read OTU mat from data file
Raw_data <- read_tsv(paste0(data_path, Seq_table), 
                        trim_ws = TRUE)

Raw_data[, 2:ncol(Raw_data)] %>% 
  t() %>% 
  as.data.frame() -> abundance_mat # convert to abundance matrix
colnames(abundance_mat) <- pull(Raw_data, "#OTU") # add sequence names

# get short names of samples
# abundance_mat %>% 
#   rownames() %>% 
#   str_remove("^Roey[0-9]{3,}-?") %>% 
#   str_split("_", simplify = T) %>% 
#   .[, 1] ->
#   short_names

# Read metadata file
Metadata <- read_csv(paste0(metadata_path, Metadata_table),
                        trim_ws = TRUE)
Metadata$Read1_file <- str_replace(Metadata$Read1_file, "(.*)_L001_R1_001_noPrimers.fastq.gz|\\.1\\.fastq.gz", "\\1")
Metadata <- Metadata[Metadata$Read1_file %in% rownames(abundance_mat), ] # remove metadata rows if the samples did not go through qual processing
# Metadata %<>% mutate(`16S copies` = replace(`16S copies`, which(`16S copies` == 0 | `16S copies` < 0), 0.001)) # replace 0 conc. with 0.001

# Order abundance_mat samples according to the metadata
sample_order <- match(rownames(abundance_mat), Metadata$Read1_file)
abundance_mat %<>% arrange(., sample_order)
rownames(abundance_mat) <- Metadata$Read1_file # needed for pyhloseq

Metadata$LibrarySize <- rowSums(abundance_mat)
Metadata <- data.frame(row.names = Metadata$Read1_file, Metadata)

# generate phyloseq object
Ps_obj <- phyloseq(otu_table(abundance_mat, taxa_are_rows = FALSE),
                        sample_data(Metadata)
                        )
Ps_obj <- filter_taxa(Ps_obj, function(x) sum(x) > 0, TRUE) # remove 0 abundance taxa
Ps_obj <- subset_samples(Ps_obj, sample_sums(Ps_obj) > 0) # remove 0 abundance samples
```

### Inspect Library Sizes
```{r Library Sizes, cache=T}
Ps_obj_df <-
  as.data.frame(sample_data(Ps_obj)) # Put sample_data into a ggplot-friendly data.frame
Ps_obj_df <- Ps_obj_df[order(Ps_obj_df$LibrarySize), ]
Ps_obj_df$Index <- seq(nrow(Ps_obj_df))
ggplot(data = Ps_obj_df, 
       aes(x = Index, y = LibrarySize, color = Control)) + 
  geom_point() +
  scale_y_log10(breaks = c(
    min(Ps_obj_df$Lib.size),
    10,
    100,
    1000,
    5000,
    10000,
    ceiling(max(Ps_obj_df$Lib.size) / 10000) * 10000
    )) + 
  scale_shape_discrete(name = "Illumina run")
summary(sample_sums(Ps_obj))
summary(taxa_sums(Ps_obj))
```

No sample has 0 counts but some controls have more sequences than any true sample. 

**This is how many reads remained per control sample after DADA2 pipeline:**
```{r}
Ps_obj %>% 
  subset_samples(., Control == "TRUE") %>% 
  sample_sums() %>% 
  sort() %>% 
  as_tibble(rownames = "Sample") %>% 
  dplyr::rename(Reads = value) %>% 
  kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```

**Summary of all non-control samples:**
```{r}
# Ps_obj <-
  # subset_samples(Ps_obj, Sample.name != names(which(sample_sums(Ps_obj) == 0)))
Ps_obj %>% 
  subset_samples(., Control == "FALSE") %>% 
  sample_sums(.) %>% 
  summary()
```

###  Identify contaminants - Frequency
Use the distribution of the frequency of each sequence feature as a function of the input DNA concentration to identify contaminants.

```{r frequency, cache=T}
Ps_obj_noNA <- subset_samples(Ps_obj, X16S.copies > 0)

contamdf.freq <-
  isContaminant(Ps_obj_noNA, method = "frequency", conc = "X16S.copies")
# print(contamdf.freq)
# How many contaminants are found?
table(contamdf.freq$contaminant)
# Which ones?
contamdf.freq %>% 
  rownames_to_column(., var = "ESV") %>% 
  filter(contaminant == TRUE) %>% 
  kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)
```

Plot the frequency of the first 20 non-contaminant ESVs against the DNA concentration, as an example.
```{r plot frequency 1, cache=T, fig.height=20, fig.width=20}
plot_frequency(Ps_obj_noNA, taxa_names(Ps_obj_noNA)[which(!contamdf.freq$contaminant)[1:20]], conc = "X16S.copies") + 
  xlab("16S copy numbers")
```

Plot the frequency of the the first 20 contaminant sequences against the DNA concentration.
```{r plot frequency 2, cache=T, fig.height=20, fig.width=20}
plot_frequency(Ps_obj_noNA, taxa_names(Ps_obj_noNA)[which(contamdf.freq$contaminant)[1:20]], conc = "X16S.copies") + 
  xlab("16S copy numbers")
```

The frequency analysis detected $`r sum(contamdf.freq$contaminant)`$ sequences as contaminants.

###  Identify contaminants - Prevalence
Use the prevalence of sequences found in the control samples (no-template controls) to identify contaminants.
```{r prevalence, cache=T}
# Ps_obj_noBioControl <- subset_samples(Ps_obj, Control != "Biological") # remove "Biological control" samples
# sample_data(Ps_obj_noBioControl)$Control <- sample_data(Ps_obj_noBioControl)$Control == "TRUE" # convert to logical
contamdf.prev <- isContaminant(Ps_obj, method = "prevalence", neg = "Control")
# How many contaminants are found?
table(contamdf.prev$contaminant)
# Which ones?
contamdf.prev %>% 
  rownames_to_column(., var = "ESV") %>% 
  filter(contaminant == TRUE) %>% 
  kable(.) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F)

# And using a more aggrssive threshold
contamdf.prev05 <- isContaminant(Ps_obj, method = "prevalence", neg = "Control", threshold = 0.5)
table(contamdf.prev05$contaminant)

# Make phyloseq object of presence-absence in negative controls
Ps_obj.pa <-
  transform_sample_counts(Ps_obj, function(abund)
    1 * (abund > 0))
Ps_obj.pa.neg <-
  prune_samples(sample_data(Ps_obj.pa)$Control == "TRUE", Ps_obj.pa)
Ps_obj.pa.pos <-
  prune_samples(sample_data(Ps_obj.pa)$Control == "FALSE", Ps_obj.pa)
# Make data.frame of prevalence in positive and negative samples
df.pa <-
  data.frame(
    pa.pos = taxa_sums(Ps_obj.pa.pos),
    pa.neg = taxa_sums(Ps_obj.pa.neg),
    contaminant = contamdf.prev$contaminant
  )
ggplot(data = df.pa, aes(x = pa.neg, y = pa.pos, color = contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")
```

The prevalence analysis detected $`r sum(contamdf.prev$contaminant)`$ sequences as contaminants.
In total $`r length(unique(c(taxa_names(Ps_obj)[which(contamdf.freq$contaminant)], taxa_names(Ps_obj)[which(contamdf.prev$contaminant)])))`$ were detected as contaminants and will be removed.

### Save contaminant sequence names
```{r save, cache=T}
c(taxa_names(Ps_obj)[which(contamdf.freq$contaminant)],
  taxa_names(Ps_obj)[which(contamdf.prev$contaminant)]) %>%
  unique() ->
  contaminants

contaminants %>% 
  data.frame() %>%
  write_csv(., 
            paste0(data_path, "decontam_contaminants.csv"), 
            col_names = FALSE)

contaminants %>% 
  setdiff(taxa_names(Ps_obj), .) %>% 
  prune_taxa(., Ps_obj) %>% 
  otu_table(.) %>% 
  as(., "matrix") %>% 
  {if(taxa_are_rows(Ps_obj)) . else t(.)} %>% 
  as_tibble() %>% 
  rownames_to_column(var = "#OTU") %>% 
  write_tsv(., 
          paste0(data_path, "DADA2.seqtab_nochim_decontam.tsv"), 
          col_names = TRUE)
```

```{r session info, cache=F}
sessionInfo()
```

## References


