---
title: "Anaerobic CUE"
subtitle: "05 Differential abundance modelling"
author: "Roey Angel"
email: "roey.angel@bc.cas.cz"
date: "`r Sys.Date()`"
bibliography: references.bib
link-citations: yes
csl: fems-microbiology-ecology.csl
output:
 rmarkdown::html_document:
  toc: true
  toc_float: true
  keep_md: true
  number_sections: false
  highlight: "pygments"
  theme: "flatly"
  dev: "png"
  df_print: "kable"
  fig_caption: true
  code_folding: "show"
---

```{r libraries, include=F}
# .libPaths(c('~/R/library', .libPaths())) # Uncomment if you have no write access to R path
library(extrafont) # ?Tools for using fonts, CRAN v0.17
library(tidyverse) # Easily Install and Load the 'Tidyverse', CRAN v1.3.0
library(scales) # Scale Functions for Visualization, CRAN v1.1.1
library(RColorBrewer)
library(ggtext)
library(magrittr) # A Forward-Pipe Operator for R, CRAN v2.0.1
library(phyloseq) # Handling and analysis of high-throughput microbiome census data, Bioconductor v1.32.0
library(speedyseq) # Faster implementations of phyloseq functions, [github::mikemc/speedyseq] v0.5.3.9001
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax
library(vegan) # Community Ecology Package, CRAN v2.5-7
library(Biostrings) # Efficient manipulation of biological strings, Bioconductor v2.56.0
library(svglite) # An 'SVG' Graphics Device, CRAN v1.2.3.2 
library(HTSSIP) # High Throughput Sequencing of Stable Isotope Probing Data Analysis, CRAN v1.4.1
library(parallel) # Support for Parallel computation in R, CRAN v4.0.3
library(ggtext) # Improved Text Rendering Support for 'ggplot2', CRAN v0.1.1
library(DESeq2) # Differential gene expression analysis based on the negative binomial distribution, Bioconductor v1.28.1
library(patchwork) 
library(ggtree)
library(ggtreeExtra)
library(visdat)
library(glue)
library(viridis)
```

```{r style settings, echo=F, message=F, warning=F, results="asis", cache=T}
options(width = 90, knitr.table.format = "html") 
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  dev = "png",
  fig.ext = "png",
  dpi = 600,
  #  fig.width = 12,
  #  fig.height = 8,
  cache.path = "05_Diff_abund_cache/",
  fig.path = "05_Diff_abund_figures/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 12
font_import(pattern = "DejaVuSans\\.", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name))
```

```{r functions, include=F}
#' scale_libraries
#'
#' Scales libraries by:
#' 1) Taking proportions
#' 2) Multiplying by a given library size of n
#' 3) Rounding
#'
#' @param ps_obj (Required) A phyloseq object
#'
#' @param n (Optional) Library size to scale to (Default: min(sample_sums(physeq)))
#'
#' @param round (Optional) Rounding method. Either "floor" (default) or "round"
#'
#' @author https://github.com/DenefLab/MicrobeMiseq/blob/master/R/miseqR.R
#' @return A phyloseq objectn ordered data frame with the columns: ASV names, taxonomic rank names and abundance
#' @usage scale_libraries(Ps_obj, n = min(sample_sums(physeq)), round = "floor")
#' @export
#'
#'

#' plot_lib_dist
#' Plot distribution of amplicon library sizes as a histogram
#'
#' @param Ps_obj (Required) A phyloseq object
#'
#' @author Roey Angel (https://github.com/roey-angel)
#' @return A ggplot object
#' @usage plot_lib_dist(Ps_obj)
#' @export

plot_lib_dist <- function(Ps_obj){
  require(ggplot2)
  require(scales)
  data.frame(sum = sample_sums(Ps_obj)) %>%
  ggplot(aes(x = sum)) +
    geom_histogram(color = "black",
                   fill = "indianred") +
    theme(panel.grid.minor = element_blank()) +
    labs(x = "Library size" , y = "Sample count") ->
    lib_dist_plot
  return(lib_dist_plot)
}

scale_libraries <- function(physeq, n = min(sample_sums(physeq)), round = "floor") {
  require(phyloseq)

  # Transform counts to n
  physeq.scale <- transform_sample_counts(physeq,
    function(x) {(n * x/sum(x))}
  )

  # Pick the rounding functions
  if (round == "floor") {
    otu_table(physeq.scale) <- floor(otu_table(physeq.scale))
  } else if (round == "round") {
    otu_table(physeq.scale) <- round(otu_table(physeq.scale), digits = 0)
  }

  # Prune taxa and return new phyloseq object
  physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)
  return(physeq.scale)
}

#' phyloseq_to_deseq2_safe
#' A wrapper for phyloseq::phyloseq_to_deseq2() but ensures that the right reference level is used
#'
#' phyloseq_to_deseq2() converts a phyloseq object to a deseq2 object. However for DESeq2 to work properly the reference level in the factor being tested needs to be the first level in the factor (R's default is alphabetical)
#'
#' @author Roey Angel (https://github.com/roey-angel)
#' @usage physeq_merge_samples(ps_obj, grouping_name = "Description")
#' @param ps (Required). A phyloseq object that has sample indices.
#'
#' @param test_condition (Required). A single character string matching a variable name in
#' the corresponding sample_data of \code{ps}.
#'
#' @param ref_level (Required). A single character string to be set as the reference level for DESeq2 (must be a factor level of test_condition)
#'
#' @return A DESeq2 object
#'
#' @seealso \code{\link{phyloseq::phyloseq_to_deseq2}}
#'
#' @export

## TODO:

phyloseq_to_deseq2_safe <- function(ps = ps_obj, test_condition = "Density.zone", ref_level = "Light") {
  require(magrittr)
  require(forcats)
  require(phyloseq)

  # critical for DESeq2 that the reference is the first level!!
  sample_data(ps)[[test_condition]] %<>% fct_relevel(., ref_level)
  expr <- as.formula(paste("~", test_condition))
  phyloseq_to_deseq2(ps, expr) %>%
    return(.)
}

mark_rare_taxa <- function(ps_obj, rank = "Phylum", rare_thresh = 0.01){
  require(dplyr)
  require(phyloseq)
  require(speedyseq)
  
  # glomerate to the "Rank" level
  ps_obj_glom <- tax_glom(ps_obj, 
                          rank, 
                          NArm = TRUE) 
  
  ps_obj_glom_rel <- transform_sample_counts(ps_obj_glom, 
                                             function(x) x / sum(x)) # transform to rel. ab.
  
  # convert to df
  ps_obj_df <- speedyseq::psmelt(ps_obj_glom_rel) # generate a df
  ps_obj_df %<>%
    mutate(Rank = as.character(!!sym(rank)))
   
  # group dataframe by Phylum, calculate sum rel. abundance
  ps_obj_df %>%
    group_by(!!sym(rank)) %>%
    summarise(Sum = sum(Abundance) / nsamples(ps_obj)) %>% 
    filter(Sum < rare_thresh) %>% # find Taxa whose mean rel. abund. is less than thresh
    pull(rank) -> 
    Rare_phyla 
  
  # change their name to "Rare"
  tax_table(ps_obj) %<>% 
    as.data.frame() %>%  
    rownames_to_column("OTU") %>% 
    mutate(across(rank, 
                  ~if_else(!!sym(rank) %in% Rare_phyla, "Rare", !!sym(rank)))) %>% 
    column_to_rownames("OTU") %>% 
    as.matrix() %>% 
    tax_table()
    return(ps_obj)
}

order_taxa <- function(ps_obj, rank = "Phylum", rel_abund = TRUE){
  require(dplyr)
  require(phyloseq)
  require(speedyseq)
  
  ps_obj %>%
    tax_glom(taxrank = rank) %>%                     # agglomerate at 'Rank' level
    {if(rel_abund) transform_sample_counts(., function(x) x / sum(x)) else .} %>% # convert to rel abundance 
    psmelt() %>%                                        # Melt to long format
    arrange(rank) %>%                                  # arrange by 'Rank'
    group_by(across(rank)) %>% 
    summarise(Abundance = sum(Abundance)) %>%
    arrange(desc(Abundance)) %>% 
    mutate(across(rank, ~factor(., levels = fct_inorder(.)))) %>%  
    mutate(across(rank, ~fct_relevel(., "Rare", after = Inf))) ->
    Taxa_order 
  
  return(Taxa_order)
}

prep_DESeq_data <- function(DESeq_results, ps_obj, sig_level = alpha_thresh, LFC = LFC_thresh, rank = "Phylum", rare_thresh = 0.01) {
  require(dplyr)
  require(magrittr)
  require(phyloseq)
  
  if (!is.list(DESeq_results)) {DESeq_results <- list(DESeq_results)} # convert to list if it is not (to support a vectorised run of DESEq2)
  
  # mark rare phyla
  ps_obj %>% mark_rare_taxa(., rank = rank, rare_thresh = rare_thresh) -> ps_obj
  
  # group dataframe by OTU, calculate median rel. abundance
  ps_obj %>%
    transform_sample_counts(., function(x) x / sum(x) * 100) %>% 
    taxa_sums(.) %>% 
    map_dbl(~(.x / nsamples(ps_obj))) %>% 
    enframe(name = "OTU", value = "Mean abundance (%)") -> 
    baseMean
  
  Taxa_order <- order_taxa(ps_obj)
  
  DESeq_results[[1]] %>% 
    as.data.frame() %>% 
    rownames_to_column(var = "OTU") %>% 
    left_join(., baseMean, by = "OTU") %>% # add mean abundance to results table
    bind_cols(., as(tax_table(ps_obj)[taxa_names(ps_obj) %in% .$OTU, ], "data.frame")) %>% # add taxnomy
    mutate(across(rank, ~factor(., levels = levels(Taxa_order$Phylum)))) %>% # order taxa by abundance
    mutate(Significance = if_else(padj < sig_level &
                            !is.na(padj) &
                            abs(lfcSE) > LFC,
                            "Pass",
                            "Fail"))  %>% 
    mutate(ymin = if_else(Significance == "Pass", log2FoldChange - lfcSE, NA_real_),
           ymax = if_else(Significance == "Pass", log2FoldChange + lfcSE, NA_real_)) -> # add error
    DESeq2plot

  return(DESeq2plot)
}

plot_DESeq <- function(DESeq_results, ps_obj, sig_level = alpha_thresh, LFC = LFC_thresh, rank = "Phylum", rare_thresh = 0.01, OTU_labels = FALSE, Y_val = "log2FoldChange", plot_title = "") {
  require(ggplot2)
  require(ggrepel)
  require(ggtext)
  
  DESeq2plot <- prep_DESeq_data(DESeq_results, ps_obj, sig_level, LFC, rank, rare_thresh)
  
  DESeq_summary <- tibble(Label = c(paste0("⬆", 
                                           sum(DESeq2plot$log2FoldChange > 0 & 
                                                 DESeq2plot$Significance == "Pass"),
                                           " (", nrow(DESeq2plot), ")")))
  
  pos <- position_jitter(width = 0.3, seed = 1)
  
  p <-
    ggplot(DESeq2plot) +
    geom_point(aes(
      x = !!sym(rank),
      y = !!sym(Y_val),
      colour = !!sym("Significance"),
      size = !!sym("Mean abundance (%)")),
      position = pos,
      alpha = 2 / 3,
      stroke = 0) +
    geom_linerange(aes(x = !!sym(rank),
                       y = !!sym(Y_val),
                       ymin = ymin,
                       ymax = ymax,
                       colour = !!sym("Significance")),
                   position = pos,
                   alpha = 1/5, 
                   show.legend = FALSE) +
    geom_text(
      data    = DESeq_summary,
      mapping = aes(x = Inf, y = Inf, label = Label),
      hjust   = 1.1,
      vjust   = 1.6
    ) +
    labs(x = "", y = "Log<sub>2</sub> fold change", title = plot_title) +
    labs(colour = paste("Significance at \n p <", sig_level), size = "Mean abundance (%)") +
    theme_grey(base_size = f_size, base_family = f_name) +
    theme(axis.text.x = element_text(angle = 45.0, vjust = 1, hjust = 1),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          axis.title.y = element_markdown()) +
    guides(colour = guide_legend(override.aes = list(size = 5))) +
    # scale_colour_manual(values = c(ggpomological:::pomological_base[[7]], ggpomological:::pomological_palette[[1]])) +
    scale_colour_manual(values = c("#2b323f", "#c03728")) +
    scale_size_continuous(name = "Mean abundance (%)",
                          range = c(2, 8),
                          breaks = c(round(seq(min(DESeq2plot$`Mean abundance (%)`), max(DESeq2plot$`Mean abundance (%)`), length.out = 5), 1)))
  
  if (OTU_labels) {
    p <- p + geom_label_repel(
      aes(x = !!sym(rank), y = !!sym(Y_val)),
      size = 6,
      label = sub("Seq_([0-9]+)", "\\1", pull(DESeq2plot[DESeq2plot$Significance == "Pass", ], "OTU")),
      position = pos,
      data = DESeq2plot[DESeq2plot$Significance == "Pass", ],
      # nudge_x = 0.4,
      colour = "#4a4a4a",
      label.size = NA, 
      alpha = 0.75, 
      # fontface = 'bold',
      box.padding = 0.80,
      point.padding = 0.5
    )
  }
  return(p)
}

plot_otus_by_density <- function(ps_obj = Ps_obj_SIP_noTime_l[[1]], 
                                 ASV2plot = filter(DESeq_res_SIP_byTime_LFC_sig_df, Site == "Certovo", Oxygen == "Anoxic"),
                                 ASV_colours = ggpomological:::pomological_palette[c(2, 4, 3, 1)],
                                 X_val = "Density..g.ml.1.", 
                                 Y_val = "Abundance", 
                                 shape_val = "Label..13C.",
                                 colour_val = "Hours",
                                 topN = 30){
  require(ggpomological)
  require(dplyr)
  require(phyloseq)
  require(speedyseq)
  
  ps_obj %>% 
    transform_sample_counts(function(x) x/sum(x) * 100) %>% 
    prune_taxa(ASV2plot$ASV, .) %>% 
    prune_taxa(unique(arrange(ASV2plot, desc(log2FoldChange))$ASV)[1:topN], .) %>% 
    # prune_taxa(names(sort(taxa_sums(.), TRUE)[1:topN]), .) %>%
    psmelt() ->
    Incorporators_df
  
  ggplot(Incorporators_df, aes(x = !!sym(X_val),
                               y = !!sym(Y_val),
                               shape = !!sym(shape_val),
                               colour = as.factor(!!sym(colour_val)))) + 
    scale_colour_manual(values = ASV_colours) +
    # scale_colour_pomological() +
    geom_point(alpha = 1/2, size = 3) + 
    geom_line() +
    facet_wrap("OTU", scales = "free_y") +
    guides(colour = guide_legend(title = "Hours"), 
           shape = guide_legend(title = "Label <sup>13</sup>C")) +
    labs(x = "Density (g ml<sup>-1</sup>)") +
    theme(legend.title = element_markdown(),
          axis.title.x = element_markdown())
  
  
}

plot_ggtree <- function(ps_obj, 
                        rank = "Phylum",
                        subrank = "Order",
                        Taxa2plot = "Actinobacteriota"){
  require(magrittr)
  require(ggplot2)
  require(ggtree)
  require(RColorBrewer)
  
  # subset (phyloseq subset_taxa doesn't support quasiquotation)
  ps_obj %>% 
    tax_table() %>%
    as("data.frame") %>%
    rownames_to_column("ASV") %>% 
    filter(., !!sym(rank) == Taxa2plot) %>% 
    column_to_rownames("ASV") %>% 
    as.matrix() %>% 
    tax_table() ->
    tax_table(ps_obj)
  
  tax_table(ps_obj) %>% 
    as("data.frame") %>% 
    pull(subrank) %>% 
    unique() %>% 
    length() ->
    n_colours

  p_tree <- ggtree(ps_obj,
                       layout = "rectangular") +
    geom_tippoint(aes(colour = !!sym(subrank)), 
                  size = 3, 
                  alpha = 3/4) +
    scale_color_manual(subrank, 
                       values = colorRampPalette(brewer.pal(11, "Spectral"))(n_colours)) +
    # geom_tiplab(hjust=-.3, size = 2) +
    geom_treescale() +
    theme_tree(legend.position = "bottom") +
    guides(colour = guide_legend(nrow = 3, 
                                 byrow = TRUE)) +
    # scale_colour_manual(values = mycolours) +
    ggtitle(Taxa2plot)
  return(p_tree)
}

plot_ggtree_heatmap <- function(ggtree, 
                                ps_obj,
                                DESeq_res_df,
                                rank = "Phylum",
                                Taxa2plot = "Actinobacteriota",
                                sample_names = "Site_Oxygen_Hours",
                                sample_colours = "black"){
  require(dplyr)
  require(ggplot2)
  require(viridis)
  
  ps_obj %>% 
    tax_table() %>%
    as("data.frame") %>%
    rownames_to_column("ASV") %>% 
    filter(., !!sym(rank) == Taxa2plot) %>% 
    column_to_rownames("ASV") %>% 
    as.matrix() %>% 
    tax_table() ->
    tax_table(ps_obj)
  
  ggtree$data %>% 
    arrange(y, label) %>% 
    pull(label) %>% 
    unique() -> tip_order
  
  DESeq_res_df %>% 
    right_join(., tibble(ASV = taxa_names(ps_obj)), 
               by  = "ASV") ->
    DDESeq_res_df_sub
  
  colours <- sample_colours
  labels <- glue("<i style='colours:{colours}'>{col_order}</i>")
  
  DDESeq_res_df_sub %>% 
    # filter(Labelled == "Labelled") %>% 
    mutate(log2FoldChange = ifelse(Labelled == "Labelled", log2FoldChange, NA_integer_)) %>% # retain values for labelled ASVs only
    select(ASV, !!sym(sample_names), log2FoldChange) %>% 
    mutate(across(ASV, ~factor(., levels = tip_order))) %>%
    complete(., !!sym(sample_names), ASV) %>% 
    ggplot(., aes(!!sym(sample_names), ASV, fill = log2FoldChange)) + 
    geom_tile(colour = "white",
              size = 0.25) +
    scale_x_discrete(expand = c(0,0),
                     name = NULL,
                     labels = labels) + 
    scale_y_discrete(expand = c(0,0)) +
    # scale_fill_gradient(low = "gray", high = "darkred") +
    scale_fill_viridis(na.value = "gray95",
                       option = "inferno") +
    guides(fill = guide_legend(title = "Log<sub>2</sub> fold change")) +
    theme(axis.text.x = element_markdown(size = 10,
                                         angle = 45, 
                                         vjust = 1, 
                                         hjust = 1),
          panel.grid.major = element_blank(),
          axis.title =  element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          legend.title = element_markdown(size = 10),
          legend.position = "bottom") ->
    HM_plot
  
  return(HM_plot)
  
}
```

## Differential abundance modelling of SIP gradients
Here we attempt to detect ASVs that were labelled with 13C in our soil incubations using differential abundance modelling.
Using DESeq2 [@love_moderated_2014] we compare the relative abundance of each ASV in the fractions where ^13^C-labelled RNA is supposed to be found (>1.795 g ml^-1^; AKA 'heavy' fractions) to the fractions where unlabelled RNA is supposed to be found (<1.795 g ml^-1^; AKA 'light' fractions)

### Setting general parameters:
```{r general parameters, cache=T}
set.seed(2021)
alpha_thresh <- 0.05
LFC_thresh <- 0.26
samples_prep_path <- "./"
data_path <- "./DADA2_pseudo/"
# Metadata_table <- "./AnCUE_Metadata_decontam.csv"
# Seq_table <- "DADA2.seqtab_nochim_decontam.tsv"
# Seq_file <- "DADA2.Seqs_decontam.fa"
Ps_file <- "Ps_obj_decontam_filt3.Rds"
Tree_file <- "./Tree/DADA2.Seqs_decontam_filtered.filtered.align.treefile"
```

### Read phyloseq object
```{r load data, cache=T}
# Load phylogenetic tree
Tree <- read_tree(paste0(data_path, Tree_file))

# load and merge  phyloseq object
readRDS(paste0(data_path, Ps_file)) %>% 
  merge_phyloseq(.,
                 phy_tree(Tree)
  ) -> Ps_obj_SIP
```

### Subset the dataset
Because the DESeq2 models will be run on each gradient separately, we need to subset This is easily done using `HTSSIP::phyloseq_subset` [@youngblut_htssip_2018]
```{r subset dataset, cache=T}
# split, ignore time points (for labelled ASV plots)
test_expr_1 <- "(Site == '${Site}' & Oxygen == '${Oxygen}' & Label..13C. == 'Unlabelled') | (Site == '${Site}'  & Oxygen == '${Oxygen}' & Label..13C. == '${Label..13C.}')"
params_1 <- get_treatment_params(Ps_obj_SIP, c("Site",
                                   "Oxygen",
                                   "Glucose",
                                   "Label..13C."),
                     "Label..13C. != 'Unlabelled'")
Ps_obj_SIP_noTime_l <- phyloseq_subset(Ps_obj_SIP, params_1, test_expr_1) 
# names(Ps_obj_SIP_noTime_l) %<>% 
#   map(., ~str_remove_all(.x, "\\s\\|\\s.*")) %>% 
#   map(., ~str_remove_all(.x, "\\(|\\)|Site == |Hours == |Oxygen == |Label..13C. == |'")) %>% 
#   map(., ~str_replace_all(.x, "([0-9]+)", "\\1 h")) 

# split, include time points (for DESeq2 modelling)
test_expr_2 <- "(Site == '${Site}' & Hours == '${Hours}' & Oxygen == '${Oxygen}' & Label..13C. == '${Label..13C.}') | (Site == '${Site}' & Hours == '${Hours}' & Oxygen == '${Oxygen}' & Label..13C. == '${Label..13C.}')"
params_2 <- get_treatment_params(Ps_obj_SIP, c("Site", 
                                   "Hours", 
                                   "Oxygen",
                                   "Glucose",
                                   "Label..13C."))

# Generate a list of subsetted phyloseq objects
Ps_obj_SIP_byTime_l <- phyloseq_subset(Ps_obj_SIP, params_2, test_expr_2) 
names(Ps_obj_SIP_byTime_l) %<>% 
  map(., ~str_remove_all(.x, "\\s\\|\\s.*")) %>% 
  map(., ~str_remove_all(.x, "\\(|\\)|Site == |Hours == |Oxygen == |Label..13C. == |'")) %>% 
  map(., ~str_replace_all(.x, "([0-9]+)", "\\1 h")) 
```

### Beta diversity analysis
Let us look first at the dissimilarity in community composition between the different fractions. If the labelling was strong enough we should see a deivation of (some of) the heavy fractions from the light ones. However, a lack of a significant deviation does not mean unsuccesful labelling because if only a small minority of the community was labelled we might not see it here (but we will, hopefully, see it using DESeq2 modelling).

```{r beta div ord joint, cache=T}
(mod1 <- adonis(vegdist(otu_table(Ps_obj_SIP), method = "horn") ~ Site * Oxygen * Hours + Lib.size,
  data = as(sample_data(Ps_obj_SIP), "data.frame"),
  permutations = 999
))

plot_lib_dist(Ps_obj_SIP)

Ps_obj_SIP %>%
  scale_libraries(round = "round") ->
  Ps_obj_SIP_scaled
  
plot_lib_dist(Ps_obj_SIP_scaled)


(mod2 <- adonis(vegdist(otu_table(Ps_obj_SIP_scaled), method = "horn") ~ Site * Oxygen * Hours + Lib.size,
  data = as(sample_data(Ps_obj_SIP_scaled), "data.frame"),
  permutations = 999
))


(mod3 <- adonis(vegdist(otu_table(Ps_obj_SIP_scaled), method = "horn") ~ Site * Oxygen * Hours * Density.zone,
  data = as(sample_data(Ps_obj_SIP_scaled), "data.frame"),
  permutations = 999
))

Site_disp <- betadisper(vegdist(otu_table(Ps_obj_SIP_scaled), method = "horn"), get_variable(Ps_obj_SIP_scaled, "Site"))
permutest(Site_disp)
plot(Site_disp)
Oxygen_disp <- betadisper(vegdist(otu_table(Ps_obj_SIP_scaled), method = "horn"), get_variable(Ps_obj_SIP_scaled, "Oxygen"))
permutest(Oxygen_disp)
plot(Oxygen_disp)
Hours_disp <- betadisper(vegdist(otu_table(Ps_obj_SIP_scaled), method = "horn"), get_variable(Ps_obj_SIP_scaled, "Hours"))
permutest(Hours_disp)
plot(Hours_disp)
Density_disp <- betadisper(vegdist(otu_table(Ps_obj_SIP_scaled), method = "horn"), get_variable(Ps_obj_SIP_scaled, "Density.zone"))
permutest(Density_disp)
plot(Density_disp)

Ord <- ordinate(Ps_obj_SIP_scaled, "CAP", "horn", 
                formula =  ~ Site * Oxygen * Hours * Density.zone)
explained <- as.numeric(format(round(eigenvals(Ord)/sum(eigenvals(Ord)) * 100, 1), nsmall = 1))
Ord_plt <- plot_ordination(Ps_obj_SIP, Ord, type = "samples", color = "Label..13C.", justDF = TRUE)

p_ord_joint <- ggplot(Ord_plt) +
  geom_point(aes(
               x = CAP1,
               y = CAP2,
               color = Label..13C.,
               size = Density..g.ml.1.,
               shape = Oxygen
             ), alpha = 2 / 3) +
  guides(colour = guide_legend(title = "Labelling"), 
         size = guide_legend(title = "Density (g ml<sup>-1</sup>)"),
         shape = guide_legend(title = "Oxygen")) +
  ggsci::scale_colour_locuszoom() +
  # scale_colour_manual(values = Gradient.colours) +
  # scale_fill_manual(values = Gradient.colours, guide = "none") +
  labs(x = sprintf("CAP1 (%s%%)", explained[1]),
  y = sprintf("CAP2 (%s%%)", explained[2])) +
  coord_fixed(ratio = sqrt(explained[2] / explained[1])) +
   theme(legend.justification = "top",
         legend.title = element_markdown(size = 11)
         ) +
  scale_size_continuous(breaks = c(seq(min(Ord_plt$Density..g.ml.1.), 
                                       max(Ord_plt$Density..g.ml.1.), 
                                       length.out = 5), 
                                   1),
                        range = c(0.1, 5)) +
  facet_grid(Site ~ Hours) +
  ggtitle("Joint analysis")
print(p_ord_joint)
```

```{r DESeq2 models by time, cache=T}
# generate a deseq object
DESeq_obj_SIP_byTime_l <- mclapply(Ps_obj_SIP_byTime_l, 
                                   function(x) {phyloseq_to_deseq2_safe(x, 
                                                                        test_condition = "Density.zone", 
                                                                        ref_level = "Light")}, 
                                   mc.cores = nrow(params_2))


# run dds pipeline
DESeq_obj_SIP_byTime_l %<>% mclapply(., 
                                     function(x) {DESeq(x, 
                                                        test = "Wald", 
                                                        fitType = "local")}, 
                                     mc.cores = nrow(params_2)) # run dds pipeline

# extract results from a DESeq analysis
DESeq_res_SIP_byTime_l  <- mclapply(DESeq_obj_SIP_byTime_l, 
                             function(x) {
                               results(x, 
                                       altHypothesis = "greater",
                                       alpha = alpha_thresh, 
                                       contrast = c("Density.zone", "Heavy", "Light"))}, # redundant if phyloseq_to_deseq2_safe() was used but doesn't hurt
                             mc.cores = nrow(params_2)) 

DESeq_res_SIP_byTime_LFC_l <- mclapply(DESeq_obj_SIP_byTime_l, 
                                     function(x) {
                                       results(x,
                                               lfcThreshold = LFC_thresh,
                                               altHypothesis = "greater",
                                               alpha = alpha_thresh,
                                               contrast = c("Density.zone", "Heavy", "Light"))}, # redundant if phyloseq_to_deseq2_safe() was used but doesn't hurt
                                     mc.cores = nrow(params_2)) # Extract results from a DESeq analysis


DESeq_res_SIP_byTime_LFC_shrink_l <- map(seq(length(DESeq_obj_SIP_byTime_l)), 
                                             ~lfcShrink(DESeq_obj_SIP_byTime_l[[.x]],
                                                         res = DESeq_res_SIP_byTime_LFC_l[[.x]],
                                                         coef = "Density.zone_Heavy_vs_Light",
                                                         type = "ashr"))
names(DESeq_res_SIP_byTime_LFC_shrink_l) <- names(DESeq_res_SIP_byTime_LFC_l)

# Compare
plotMA(DESeq_res_SIP_byTime_l[[2]], ylim = c(-2,2))
plotMA(DESeq_res_SIP_byTime_LFC_l[[2]], ylim = c(-2,2))
plotMA(DESeq_res_SIP_byTime_LFC_shrink_l[[2]], ylim = c(-2,2))

# summarise results (lfcShrink doesn't change the values)
# map2(DESeq_res_SIP_byTime_l, print(names(DESeq_res_SIP_byTime_l)), ~summary(.x)) # summarise results
for (i in seq(1, length(DESeq_res_SIP_byTime_l))) { # didn't manage with map
  print(names(DESeq_res_SIP_byTime_l[i]))
  summary(DESeq_res_SIP_byTime_l[[i]])
}

for (i in seq(1, length(DESeq_res_SIP_byTime_LFC_l))) { # didn't manage with map
  print(names(DESeq_res_SIP_byTime_LFC_l[i]))
  summary(DESeq_res_SIP_byTime_LFC_l[[i]])
}

# Store labelled OTUs and save them to a file
# DESeq_res_SIP_byTime_l %>% 
#   map(., ~subset(.x, padj < alpha_thresh & log2FoldChange > LFC_thresh)) %>% 
#   map(., ~as.data.frame(.x)) %>% 
#   map(., ~rownames_to_column(.x, "ASV")) %>% 
#   bind_rows(., .id = "Comparison") %>% 
#   arrange(Comparison, desc(baseMean)) %T>% 
#   write_csv(., file = "DESeq2_byTime_a-0.05.txt") ->
#   DESeq_res_SIP_byTime_df

DESeq_res_SIP_byTime_LFC_l %>% 
  map(., ~subset(.x, padj < alpha_thresh & log2FoldChange > LFC_thresh)) %>% 
  map(., ~as.data.frame(.x)) %>% 
  map(., ~rownames_to_column(.x, "ASV")) %>% 
  bind_rows(., .id = "Comparison") %>% 
  arrange(Comparison, desc(baseMean)) %>% 
  separate(., "Comparison" ,c("Site","Hours", "Oxygen", "Label"), sep = " & ") %T>% 
  write_csv(., file = "DESeq2_byTime_a-0.05_LFC0-322.txt") ->
  DESeq_res_SIP_byTime_LFC_sig_df
```

```{r vis DES res, cache=T}
DESeq_res_SIP_byTime_LFC_sig_df %>% 
  get_variable() %>% 
  select_if(is.numeric) %>% 
  vis_value()

DESeq_res_SIP_byTime_LFC_sig_df %>% 
  get_variable() %>% 
  select_if(is.numeric) %>% 
  vis_cor()
```

```{r plot DESeq2 models, cache=T}
# ps_obj <- Ps_obj_SIP
# DESeq_results <- DESeq_res_SIP_byTime_LFC0.322_l[9]
# plot_DESeq(DESeq_results, ps_obj, plot_title = names(DESeq_results))

DESeq_plots <- lapply(seq(length(DESeq_res_SIP_byTime_LFC_shrink_l)), 
                        function(x) {plot_DESeq(DESeq_res_SIP_byTime_LFC_shrink_l[x],  
                                                Ps_obj_SIP, plot_title = names(DESeq_res_SIP_byTime_LFC_shrink_l[x]))})


(DESeq_plots[[6]] + 
    theme(legend.position = "none") +
    theme(axis.text.x = element_blank())) +
  (DESeq_plots[[1]] + 
     theme(legend.position = "none", 
           axis.text.x = element_blank(), 
           axis.title.y = element_blank())) +
  (DESeq_plots[[7]] + 
     theme(legend.position = "none",
           axis.text.x = element_blank())) +
  (DESeq_plots[[3]] + 
     theme(legend.position = "none", 
           axis.text.x = element_blank(), 
           axis.title.y = element_blank())) +
  (DESeq_plots[[8]] + 
     theme(legend.position = "none") +
     theme(legend.position = "none",
           axis.text.x = element_blank())) +
  (DESeq_plots[[4]] + 
     theme(legend.position = "none") +
     theme(legend.position = "none", 
           axis.text.x = element_blank(), 
           axis.title.y = element_blank()) +
     ylim(NA, 5)) +
  (DESeq_plots[[9]] + 
     theme(legend.position = "none",
           axis.text.x = element_blank())) +
  (DESeq_plots[[2]] + 
     theme(legend.position = "none", 
           axis.text.x = element_blank(), 
           axis.title.y = element_blank())) +
  (DESeq_plots[[10]] + 
     theme(legend.position = "none")) +
  (DESeq_plots[[5]] + 
     theme(legend.position = "none", 
           axis.title.y = element_blank())) + 
  plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = 'bottom')

  
(DESeq_plots[[16]] + 
    theme(legend.position = "none") +
    theme(axis.text.x = element_blank())) +
  (DESeq_plots[[11]] + 
     theme(legend.position = "none", 
           axis.text.x = element_blank(), 
           axis.title.y = element_blank())) +
  (DESeq_plots[[17]] + 
     theme(legend.position = "none",
           axis.text.x = element_blank())) +
  (DESeq_plots[[13]] + 
     theme(legend.position = "none", 
           axis.text.x = element_blank(), 
           axis.title.y = element_blank())) +
  (DESeq_plots[[18]] + 
     theme(legend.position = "none") +
     theme(legend.position = "none",
           axis.text.x = element_blank())) +
  (DESeq_plots[[14]] + 
     theme(legend.position = "none") +
     theme(legend.position = "none", 
           axis.text.x = element_blank(), 
           axis.title.y = element_blank())) +
  (DESeq_plots[[19]] + 
     theme(legend.position = "none",
           axis.text.x = element_blank())) +
  (DESeq_plots[[12]] + 
     theme(legend.position = "none", 
           axis.text.x = element_blank(), 
           axis.title.y = element_blank()) +
     ylim(-3, NA)) +
  (DESeq_plots[[20]] + 
     theme(legend.position = "none")) +
  (DESeq_plots[[15]] + 
     theme(legend.position = "none", 
           axis.title.y = element_blank())) + 
  plot_layout(ncol = 2, guides = "collect") & 
  theme(legend.position = 'bottom')
```

```{r Plot incorporators, cache=T}
plot_otus_by_density(ps_obj = Ps_obj_SIP_noTime_l[[2]], 
                     ASV2plot = filter(DESeq_res_SIP_byTime_LFC_sig_df, Site == "Certovo", Oxygen == "Oxic"))
plot_otus_by_density(ps_obj = Ps_obj_SIP_noTime_l[[1]], 
                     ASV2plot = filter(DESeq_res_SIP_byTime_LFC_sig_df, Site == "Certovo", Oxygen == "Anoxic"))
plot_otus_by_density(ps_obj = Ps_obj_SIP_noTime_l[[4]], 
                     ASV2plot = filter(DESeq_res_SIP_byTime_LFC_sig_df, Site == "Plesne", Oxygen == "Oxic"))
plot_otus_by_density(ps_obj = Ps_obj_SIP_noTime_l[[3]],
                     ASV2plot = filter(DESeq_res_SIP_byTime_LFC_sig_df, Site == "Plesne", Oxygen == "Anoxic"))
```

```{r Plot trees, cache=T}
c("Certovo Oxic 12 h", "Certovo Oxic 24 h", "Certovo Oxic 48 h", "Certovo Oxic 72 h", "Certovo Anoxic 12 h", "Certovo Anoxic 24 h", "Certovo Anoxic 48 h", "Certovo Anoxic 216 h", "Plesne Oxic 12 h", "Plesne Oxic 24 h", "Plesne Oxic 48 h", "Plesne Oxic 72 h", "Plesne Anoxic 12 h",  "Plesne Anoxic 24 h",  "Plesne Anoxic 48 h",  "Plesne Anoxic 216 h" )  ->
  col_order

DESeq_res_SIP_byTime_LFC_l %>% 
  map(., ~as.data.frame(.x)) %>% 
  map(., ~rownames_to_column(.x, "ASV")) %>% 
  bind_rows(., .id = "Comparison") %>% 
  filter(str_detect(Comparison, "Labelled")) %>% # remove unlabelled samples [c(-5, -10, -15, -20)]
  mutate(Labelled = ifelse(padj < alpha_thresh & log2FoldChange > LFC_thresh, "Labelled", "Unlabelled")) %>% 
  # arrange(Comparison, desc(baseMean)) %>% 
  separate(., "Comparison" ,c("Site","Hours", "Oxygen", "Label"), sep = " & ") %>% 
  mutate(Site_Oxygen_Hours = paste(Site, Oxygen, Hours)) %>% 
  mutate(across(Site_Oxygen_Hours, ~factor(., levels = col_order))) ->
  # mutate(Site_oxygen = paste(Site, Oxygen)) ->
  DESeq_res_SIP_byTime_all_df

# Summarise number of labelled and unlabelled ASVs
DESeq_res_SIP_byTime_all_df %>% 
  group_by(Labelled) %>% 
  summarise(n = n()) 

# detect taxa with NA from DESeq analysis
DESeq_res_SIP_byTime_all_df %<>% 
  filter(!is.na(Labelled)) #%>% 
  # pull(Labelled) -> 
  # bad_seqs

# remove NA taxa from PS obj
Ps_obj_SIP %>% 
  prune_taxa(DESeq_res_SIP_byTime_all_df$ASV, .) -> 
  Ps_obj_SIP4tree_plot

p_actino_t <- plot_ggtree(Ps_obj_SIP4tree_plot, 
                          rank = "Class",
                          subrank = "Order",
                          Taxa2plot = "Actinobacteria")
p_actino_HM <- plot_ggtree_heatmap(p_actino_t, 
                                 Ps_obj_SIP4tree_plot,
                                 DESeq_res_SIP_byTime_all_df, 
                                 rank = "Class",
                                 Taxa2plot = "Actinobacteria",
                                 sample_names = "Site_Oxygen_Hours",
                                 sample_colours = rep(brewer.pal(n = 11, 
                                                                 "RdYlGn")[c(1, 3, 11, 9)],
                                                      each = 4))
p_actino_t + p_actino_HM

p_Aproteo_t <- plot_ggtree(Ps_obj_SIP4tree_plot, 
                          rank = "Class",
                          subrank = "Order",
                          Taxa2plot = "Alphaproteobacteria")
p_Aproteo_HM <- plot_ggtree_heatmap(p_Aproteo_t, 
                                 Ps_obj_SIP4tree_plot,
                                 DESeq_res_SIP_byTime_all_df, 
                                 rank = "Class",
                                 Taxa2plot = "Alphaproteobacteria",
                                 sample_names = "Site_Oxygen_Hours",
                                 sample_colours = rep(brewer.pal(n = 11, 
                                                                 "RdYlGn")[c(1, 3, 11, 9)],
                                                      each = 4))
p_Aproteo_t + p_Aproteo_HM

p_Gproteo_t <- plot_ggtree(Ps_obj_SIP4tree_plot, 
                          rank = "Class",
                          subrank = "Order",
                          Taxa2plot = "Gammaproteobacteria")
p_Gproteo_HM <- plot_ggtree_heatmap(p_Gproteo_t, 
                                 Ps_obj_SIP4tree_plot,
                                 DESeq_res_SIP_byTime_all_df, 
                                 rank = "Class",
                                 Taxa2plot = "Gammaproteobacteria",
                                 sample_names = "Site_Oxygen_Hours",
                                 sample_colours = rep(brewer.pal(n = 11, 
                                                                 "RdYlGn")[c(1, 3, 11, 9)],
                                                      each = 4))
p_Gproteo_t + p_Gproteo_HM

p_acido_t <- plot_ggtree(Ps_obj_SIP4tree_plot, 
                          rank = "Phylum",
                          subrank = "Class",
                          Taxa2plot = "Acidobacteriota")
p_acido_HM <- plot_ggtree_heatmap(p_acido_t, 
                                 Ps_obj_SIP4tree_plot,
                                 DESeq_res_SIP_byTime_all_df, 
                                 rank = "Phylum",
                                 Taxa2plot = "Acidobacteriota",
                                 sample_names = "Site_Oxygen_Hours",
                                 sample_colours = rep(brewer.pal(n = 11, 
                                                                 "RdYlGn")[c(1, 3, 11, 9)],
                                                      each = 4))
p_acido_t + p_acido_HM

p_verruc_t <- plot_ggtree(Ps_obj_SIP4tree_plot, 
                          rank = "Phylum",
                          subrank = "Class",
                          Taxa2plot = "Verrucomicrobiota")
p_verruc_HM <- plot_ggtree_heatmap(p_verruc_t, 
                                 Ps_obj_SIP4tree_plot,
                                 DESeq_res_SIP_byTime_all_df, 
                                 rank = "Phylum",
                                 Taxa2plot = "Verrucomicrobiota",
                                 sample_names = "Site_Oxygen_Hours",
                                 sample_colours = rep(brewer.pal(n = 11, 
                                                                 "RdYlGn")[c(1, 3, 11, 9)],
                                                      each = 4))
p_verruc_t + p_verruc_HM

p_bacteroi_t <- plot_ggtree(Ps_obj_SIP4tree_plot, 
                          rank = "Phylum",
                          subrank = "Class",
                          Taxa2plot = "Bacteroidota")
p_bacteroi_HM <- plot_ggtree_heatmap(p_bacteroi_t, 
                                 Ps_obj_SIP4tree_plot,
                                 DESeq_res_SIP_byTime_all_df, 
                                 rank = "Phylum",
                                 Taxa2plot = "Bacteroidota",
                                 sample_names = "Site_Oxygen_Hours",
                                 sample_colours = rep(brewer.pal(n = 11, 
                                                                 "RdYlGn")[c(1, 3, 11, 9)],
                                                      each = 4))
p_bacteroi_t + p_bacteroi_HM

p_firmi_t <- plot_ggtree(Ps_obj_SIP4tree_plot, 
                          rank = "Phylum",
                          subrank = "Class",
                          Taxa2plot = "Firmicutes")
p_firmi_HM <- plot_ggtree_heatmap(p_firmi_t, 
                                 Ps_obj_SIP4tree_plot,
                                 DESeq_res_SIP_byTime_all_df, 
                                 rank = "Phylum",
                                 Taxa2plot = "Firmicutes",
                                 sample_names = "Site_Oxygen_Hours",
                                 sample_colours = rep(brewer.pal(n = 11, 
                                                                 "RdYlGn")[c(1, 3, 11, 9)],
                                                      each = 4))
p_firmi_t + p_firmi_HM

```
```{r, eval=F}
c("Certovo Oxic 12 h", "Certovo Oxic 24 h", "Certovo Oxic 48 h", "Certovo Oxic 72 h", "Certovo Anoxic 12 h", "Certovo Anoxic 24 h", "Certovo Anoxic 48 h", "Certovo Anoxic 216 h", "Plesne Oxic 12 h", "Plesne Oxic 24 h", "Plesne Oxic 48 h", "Plesne Oxic 72 h", "Plesne Anoxic 12 h",  "Plesne Anoxic 24 h",  "Plesne Anoxic 48 h",  "Plesne Anoxic 216 h" ) %>% 
 map_dfr(~tibble(!!.x := double())) ->
  col_order

DESeq_res_SIP_byTime_all_df_actino %>% 
  filter(Labelled == "Labelled") %>% 
  select(ASV, Site, Oxygen, Hours, log2FoldChange) %>% 
  pivot_wider(names_from = c(Site, Oxygen, Hours), 
              values_from = log2FoldChange,
              names_sep = " ") %>% 
  bind_rows(col_order, .) %>% 
  # add_column(., !!!heatmap_blnk_df[setdiff(names(heatmap_blnk_df), names(.))])
  # set_names() %>% 
  # select(c("ASV", "Certovo Oxic_12 h", "Certovo Oxic_24 h", "Certovo Oxic_48 h", "Certovo Oxic_72 h", "Certovo Anoxic_12 h", "Certovo Anoxic_24 h", "Certovo Anoxic_48 h", "Certovo Anoxic_216 h", "Plesne Oxic_12 h", "Plesne Oxic_24 h", "Plesne Oxic_48 h", "Plesne Oxic_72 h", "Plesne Anoxic_12 h",  "Plesne Anoxic_24 h",  "Plesne Anoxic_48 h",  "Plesne Anoxic_216 h" )) %>%
  # modify_if(is.double, ~replace_na(.x, "")) %>%
  column_to_rownames("ASV") %>% 
  set_colnames(labels) ->
  DESeq_res_SIP_byTime_all_df_actino_hm 

gheatmap(
  p = p_actino_t, 
  data = DESeq_res_SIP_byTime_all_df_actino_hm,
  offset = 0, width = 1.6, font.size = 3, 
  low = "grey", 
  high = "darkred",
  color = "grey95",
  colnames_position = "bottom",
  colnames_offset_x = -0.05,
  colnames_offset_y = 1,
  colnames_angle = 25, 
  hjust = 0,
  legend_title = "L2FC") +
  theme_tree(legend.position = "bottom") +
  theme(
    axis.text.x = element_markdown(color = "red", size = 11)
  ) +
  guides(colour = guide_legend(nrow = 3, 
                               byrow = TRUE))


p <- ggtree(Actino_ps) + 
  # geom_treescale() + 
  geom_tiplab(size=2) +
  # geom_treescale(x=2008, y=1, offset=2) 
gheatmap(p, DESeq_res_SIP_byTime_all_df_actino_hm) +
  # scale_fill_brewer()
  scale_fill_manual(breaks=c("Unlabelled", "Labelled"), 
                    values=c("firebrick", "darkgreen"), name="DESeq_res_SIP_byTime_all_df_actino_hm")


gheatmap(p_actino_t, 
         DESeq_res_SIP_byTime_all_df_actino_hm, offset = 10, color = NULL, 
         colnames_position = "top", 
         colnames_angle = 90, colnames_offset_y = 1, 
         hjust = 0, font.size = 2)


+
  scale_fill_manual(values = heatmap.colours, breaks=0:15)




library(ggnewscale)

# p_actino_t %<+%  DESeq_res_SIP_byTime_all_df_actino# add data
# p_actino_t + 
#         new_scale_colour() + 
#   geom_fruit(
#           data = DESeq_res_SIP_byTime_all_df_actino,
#           geom = geom_point,
#           mapping = aes(y = ASV, x = Site_oxygen, fill = Labelled),
#           offset=0.08,   # The distance between layers, default is 0.03 of x range of tree.
#           pwidth=0.25, # width of the layer, default is 0.2 of x range of tree.
#           axis.params=list(                
#                         axis="x",  # add axis text of the layer.
#                         text.angle=-45, # the text size of axis.
#                         hjust=0  # adust the horizontal position of text of axis.
#                       )
#       ) 

beast_file <- system.file("examples/MCC_FluA_H3.tree", package="ggtree")
beast_tree <- read.beast(beast_file)

genotype_file <- system.file("examples/Genotype.txt", package="ggtree")
genotype <- read.table(genotype_file, sep="\t", stringsAsFactor=F)
colnames(genotype) <- sub("\\.$", "", colnames(genotype))
p <- ggtree(beast_tree, mrsd="2013-01-01") + 
  geom_treescale(x=2008, y=1, offset=2) + 
  geom_tiplab(size=2)
gheatmap(p, genotype, offset=5, width=0.5, font.size=3, 
         colnames_angle=-45, hjust=0) +
  scale_fill_manual(breaks=c("HuH3N2", "pdm", "trig"), 
                    values=c("steelblue", "firebrick", "darkgreen"), name="genotype")


```

```{r colophon, eval=T}
sessioninfo::session_info() %>%
  details::details(
    summary = 'Current session info',
    open    = TRUE
 )
```

## References

