---
title: "Anaerobic CUE"
subtitle: "Plotting of C-partitioning trends and other chemical data"
description: "V1.0"   
author: 
  name: "Roey Angel"
  date: "`r Sys.Date()`"
  affiliation: "Biology Centre CAS"
  email: "roey.angel@bc.cas.cz"

csl: fems-microbiology-ecology.csl
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    keep_md: true
    number_sections: false
    highlight: "pygments"
    theme: "flatly"
    df_print: "kable"
    fig_caption: true
    code_folding: "show"
---

```{r libraries, include=F}
# .libPaths(c('~/R/library', .libPaths())) # Uncomment if you have no write access to R path
library(tidyverse)
library(scales)
library(extrafont)
library(see)
library(ggsci)
library(ggtext)
library(ggrepel)
library(patchwork)
library(ggpomological)
library(paletteer)
```

```{r style settings, echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
options(width = 90, knitr.table.format = "html") 
knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  #dev = "ragg_png",
  #fig.ext = "png",
  dev = c("svglite", "ragg_png"),
  dev.args = list(svglite = list(bg = 'white', fix_text_size = FALSE), ragg_png = list(bg = 'white')),
  dpi = 300,
  cache.path = "C_part_cache/",
  fig.path = "C_part_figures/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 14
font_import(pattern = "DejaVuSans\\.", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name)) 
pom5 <- ggpomological:::pomological_palette[c(2, 1, 3, 4, 5)]
pom2 <- c(ggpomological:::pomological_base[[7]], ggpomological:::pomological_palette[[1]])
```

```{r functions, include=F}

```

## Respiration and biomass
```{r load resp_biomass data, cache=T}
read_delim("./Chem_isotope_data/Resp_BMC.csv", delim = ";" ) %>%
  mutate(across(c("Aeration", 
                  "Site"), ~factor(.))) %>% 
  mutate(`Site:Aeration` = paste(Site, Aeration)) %>% 
  mutate(across(Site, ~fct_relevel(., "Plešné"))) %>% 
  mutate(across(Aeration, ~fct_relevel(., "Oxic"))) %>% 
  identity() ->
  Resp_MB
```

```{r plot fumigation, cache=T}
tot_fumigation <- ggplot(data = Resp_MB, aes(x = `Time (h)`, y = `C-MB (µmol(12+13))`, colour = Site, shape = Aeration)) +
  geom_point2(size = 4, alpha = 2/5) + 
  stat_summary(fun = mean, 
               geom = "point", 
               size = 7, 
               alpha = 3/4) +
  stat_summary(fun.data = mean_se,  
               geom = "errorbar",
               width = 4,
               alpha = 3/4) +
  stat_summary(fun = mean, geom = "line",alpha = 3/4) +
  # geom_boxplot(aes(x = `Time (h)`, group = `Site:Aeration`), color = "gray60", outlier.alpha = 0) +
  scale_colour_manual(values = pom5) +
  # scale_color_uchicago() +
  labs(y = "Microbial biomass (\u03bcmol <sup>12</sup>C + <sup>13</sup>C)") +
  theme(
    axis.title.y = element_markdown()
  )
```

```{r plot respiration, cache=T}
respiration <- ggplot(data = Resp_MB, aes(x = `Time (h)`, y = `Total resp (12C+13C umol/g)`, colour = Site, shape = Aeration)) +
  geom_point2(size = 4, alpha = 2/5) + 
  stat_summary(fun = mean, 
               geom = "point", 
               size = 7, 
               alpha = 3/4) +
  stat_summary(fun.data = mean_se,  
               geom = "errorbar",
               width = 4,
               alpha = 3/4) +
  stat_summary(fun = mean, geom = "line",alpha = 3/4) +
  # geom_boxplot(aes(x = `Time (h)`, group = `Site:Aeration`), colour = "gray60", outlier.alpha = 0) +
  scale_colour_manual(values = pom5) +
  # scale_color_uchicago() +
  labs(y = "CO<sub>2</sub> (\u03bcmol [<sup>12</sup>C + <sup>13</sup>C] g<sup>-1</sup>)") +
  theme(
    axis.title.y = element_markdown()
  )
```

```{r combine resp. and biomass, cache=F, fig.width=18, fig.asp=.42}
respiration + 
  theme(legend.position = "none") + 
  tot_fumigation + theme(legend.position = c(0.9, 0.84)) +
  plot_annotation(tag_levels = 'A')
```

## Glucose C partitioning
```{r load C partitioning data, cache=T}
read_delim("./Chem_isotope_data/C_part.csv", delim = ";" ) %>%
  mutate(across(c("Aeration", 
                  "Site",
                  "Time (h)"), ~factor(.))) %>% 
  dplyr::select(!starts_with("C from")) %>% 
  dplyr::select(!`WEC from CH2O % + CHO`) %>% 
  rename_with(~ gsub(" %", "", .x)) %>% 
  rename_with(~ gsub("CH2O", "glucose", .x)) %>% 
  rename(Glucose = glucose) %>% 
  rename(`CO<sub>2</sub> from glucose` = `CO2 from glucose`) %>% 
  mutate(across(Site, ~fct_relevel(., "Plešné"))) %>% 
  mutate(across(Aeration, ~fct_relevel(., "Oxic"))) %>% 
  pivot_longer(cols = c(-"Aeration", -"Site", -"Time (h)"), names_to = "C pool", values_to = "Fraction (%)") %>% 
  mutate(across(`C pool`, ~fct_relevel(., "Glucose"))) %>%
  identity() ->
  C_part
```

```{r plot C_part, cache=T}
C_part_bar <- ggplot(data = C_part, aes(x = `Time (h)`, y = `Fraction (%)`, fill = `C pool`)) +
  geom_bar(position = "stack", stat = "identity") +
  facet_grid(Site ~ Aeration, scales = "free_x") +
  theme(legend.text = element_markdown()) +
  scale_fill_manual(values = pom5)
# scale_fill_pomological()
C_part_bar
```
**Comments:**
1. The original figure had a red line at about 110% and the caption said "Red line denotes amount of added C and the difference between columns and line show proportion of C lost into non extractable C pool" But why not at 100%?

2. We might get criticised for not providing errors here, but because these are all derived values one would have to propagate errors here and that's time consuming

## C pools total
```{r load C bulk data, cache=T}
read_delim("./Chem_isotope_data/C_total.csv", delim = ";" ) %>%
  mutate(across(c("Aeration", 
                  "Site",
                  "Time (h)"), ~factor(.))) %>% 
  rename_with(~gsub("CO2", "CO<sub>2</sub>", .x)) %>% 
  rename_with(~gsub("K2SO4", "K<sub>2</sub>SO<sub>4</sub>", .x)) %>% 
  rename_with(~gsub(" \\(µmol\\[12C\\+13C\\] g-1\\)", "", .x)) %>%
  mutate(across(Site, ~fct_relevel(., "Plešné"))) %>% 
  mutate(across(Aeration, ~fct_relevel(., "Oxic"))) %>% 
  pivot_longer(cols = c(-"Aeration", -"Site", -"Time (h)"), names_to = "C pool", values_to = "C (µmol[12C+13C] g-1)") %>% 
  mutate(across(`C pool`, ~fct_relevel(., "Glucose"))) %>%
  replace(is.na(.), 0) %>% 
  identity() ->
  C_bulk
```

```{r plot C_bulk, cache=T}
C_bulk %>% # the names of the new data frame and the data frame to be summarised
  group_by(Site, Aeration,  `Time (h)`, `C pool`) %>%   # the grouping variable
  summarise(`Mean C (µmol[12C+13C] g-1)` = mean(`C (µmol[12C+13C] g-1)`),
            sd_C = sd(`C (µmol[12C+13C] g-1)`), # calculates the standard deviation of each group
            n_C = n(),  # calculates the sample size per group
            SE_C = sd(`C (µmol[12C+13C] g-1)`)/sqrt(n())) %>% 
  arrange(desc(`C pool`)) %>% 
  mutate(stackedY = cumsum(`Mean C (µmol[12C+13C] g-1)`)) -> 
  C_bulk_means

C_bulk_bar <- ggplot(data = C_bulk_means, aes(x = `Time (h)`,
                                              y = `Mean C (µmol[12C+13C] g-1)`,
                                              fill = `C pool`)) +
  geom_bar(stat = "identity") +
  geom_errorbar(inherit.aes = FALSE,
                aes(x = `Time (h)`,
                    ymin = stackedY - SE_C,
                    ymax = stackedY + SE_C),
                width = 0.2,
                alpha = 3/4) +
  facet_grid(Site ~ Aeration, scales = "free_x") +
  labs(y = "CO<sub>2</sub> (\u03bcmol [<sup>12</sup>C + <sup>13</sup>C] g<sup>-1</sup>)") +
  theme(legend.text = element_markdown(),
        axis.title.y = element_markdown()) +
  scale_fill_manual(values = pom5)
C_bulk_bar
```

**Comments:**
1. Why is this graph needed?
2. Missing values but they appear in the original graph (means)

## Organic acids
```{r load org C, cache=T}
read_delim("./Chem_isotope_data/OA.csv", delim = ";" ) %>%
  mutate(across(c("Aeration", 
                  "Site",
                  "Time (h)"), ~factor(.))) %>% 
  mutate(across(Site, ~fct_relevel(., "Plešné"))) %>% 
  mutate(across(Aeration, ~fct_relevel(., "Oxic"))) %>% 
  pivot_longer(cols = c(-"Number", -"Code", -"Aeration", -"Site", -"Time (h)"), names_to = "Organic acid", values_to = "C (µmol g-1)") %>% 
  replace(is.na(.), 0) %>% 
  identity() ->
  Organics
```

```{r plot org C, cache=T}
Organics %>% # the names of the new data frame and the data frame to be summarised
  filter(`Organic acid` != "OAtot") %>% 
  group_by(Site, Aeration,  `Time (h)`, `Organic acid`) %>%   # the grouping variable
  summarise(`Mean C (µmol g-1)` = mean(`C (µmol g-1)`),
            sd_C = sd(`C (µmol g-1)`), # calculates the standard deviation of each group
            n_C = n(),  # calculates the sample size per group
            SE_C = sd(`C (µmol g-1)`)/sqrt(n())) %>% 
  arrange(desc(`Organic acid`)) %>% 
  mutate(stackedY = cumsum(`Mean C (µmol g-1)`)) -> 
  Organics_means

Organics_bar <- ggplot(data = Organics_means, aes(x = `Time (h)`,
                                                  y = `Mean C (µmol g-1)`,
                                                  fill = `Organic acid`)) +
  geom_bar(stat = "identity") +
  geom_errorbar(inherit.aes = FALSE,
                aes(x = `Time (h)`,
                    ymin = stackedY - SE_C,
                    ymax = stackedY + SE_C),
                width = 0.2,
                alpha = 2/4) +
  facet_grid(Site ~ Aeration, scales = "free_x") +
  labs(y = "C (&mu;mol g<sup>-1</sup>)") +
  theme(axis.title.y = element_markdown()) +
  # scale_fill_scico_d(palette = "fes") 
  scale_fill_manual(values = (colorRampPalette(ggpomological:::pomological_palette)(12)))
Organics_bar
```

```{r combine_resp_biomass_Cpart_OA, cache=F, fig.width=18, fig.asp=.6}
C_part_e_bar <- readRDS(file = "C_part_e_bar.Rds") # from C_partitioning.Rmd
(C_part_e_bar + theme(legend.position = "bottom") + 
    Organics_bar) /
  (respiration + 
     theme(legend.position = "none") + 
     tot_fumigation + theme(legend.position = "top")) +
  
  
  plot_annotation(tag_levels = 'A')
```

## Oxygen
```{r oxygen, cache=T, fig.width=8, fig.asp=.6}
read_delim("./Chem_isotope_data/O2.csv", delim = ";" ) %>%
  mutate(across(c("Site"),
                ~factor(.))) %>% 
  mutate(across(Site, ~fct_relevel(., "Plešné"))) %>% 
  mutate(`O2 (%)` = `O2 in the headspace (ppmv)`/10000) %>% 
  identity() ->
  Oxygen

ggplot(data = Oxygen, aes(x = `Time (h)`, 
                          y = `O2 (%)`,
                          colour = Site)) + 
  geom_point2(size = 4, alpha = 1/3) +
  geom_smooth(method = "lm", alpha = .15, aes(fill = Site)) +
  labs(y = "Headspace O<sub>2</sub> (%)") +
  theme(legend.text = element_markdown(),
        axis.title.y = element_markdown()) +
  scale_y_continuous(limits=c(0, 22)) + 
  scale_colour_pomological() +
  scale_fill_pomological()
```

## Biomass change
```{r load biomass, cache=T}
read_delim("./Chem_isotope_data/New_biomass.csv", delim = ";" ) %>%
  filter(`Time (h)` != 0.5) %>%  
  mutate(across(c("Aeration", 
                  "Site",
                  `Time (h)`), ~factor(.))) %>% 
  mutate(across(Site, ~fct_relevel(., "Plešné"))) %>% 
  mutate(across(Aeration, ~fct_relevel(., "Oxic"))) %>% 
  mutate(across(Method, ~recode(., d13C = "&delta;<sup>13</sup>C", d2H = "&delta;<sup>2</sup>H", Fum = "Chloroform fumigation"))) ->
  Biomass
```

```{r plot biomass, cache=T}
Biomass_bar <- ggplot(data = Biomass, aes(x = `Time (h)`,
                                          y = `New biomass`,
                                          fill = `Method`)) +
  geom_bar(position = position_dodge(),  
           stat = "identity") +
  geom_errorbar(aes(group = `Method`,
                    ymin = `New biomass` - SEM,
                    ymax = `New biomass` + SEM),
                position=position_dodge(0.9),
                width = 0.2,
                alpha = 2/4) +
  facet_grid(Aeration ~ Site, scales = "free") +
  labs(y = "New biomass (&mu;mol g<sup>-1</sup>)") +
  theme(axis.title.y = element_markdown(),
        legend.text = element_markdown()) +
  scale_fill_manual(values = rev(pom5[1:3])) +
  scale_y_continuous(
    trans = scales::pseudo_log_trans(sigma = 10^(-4)), 
    breaks = trans_breaks("log10", function(x) 10^x,)(c(0.01, 1, 10, 100)),
    labels = trans_format("log10", math_format(10^.x)))
# scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
#               labels = trans_format("log10", math_format(10^.x)))
Biomass_bar
```

## CUE over time

```{r cue time, cache=T}
read_delim("./Chem_isotope_data/CUE_time.csv", delim = ";" ) %>% 
  mutate(across(c("Aeration",
                  "Site"), ~factor(.))) %>%
  mutate(across(Site, ~fct_relevel(., "Plešné"))) %>%
  mutate(across(Aeration, ~fct_relevel(., "Oxic"))) ->
  CUE_time
```

```{r cue time plot, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.asp=0.5}
CUE_time_plot <- ggplot(data = CUE_time, aes(x = `Time (h)`, y = Value, colour = Site, shape = Measurement, label = Stat)) +
  geom_point(aes(fill = Site), size = 4, alpha = 4/5) + 
  geom_line(alpha = 3/5) +
  # scale_shape_manual(values=c("O" = 21,"A" = 24, "OA" = 1, "AA" = 2)) +
  # scale_shape_manual(values=c("PL" = 21,"CT" = 24, "PLA" = 1, "CTA" = 2)) +
  scale_fill_manual(values = pom5) +
  scale_colour_manual(values = pom5) +

  # scale_color_uchicago() +
  labs(y = "Carbon use effeciency") +
  theme(
    axis.title.y = element_markdown()
  ) + 
  # geom_text_repel() +
  facet_wrap(vars(Aeration))
CUE_time_plot
```


```{r colophon, eval=T}
sessioninfo::session_info() %>%
  details::details(
    summary = 'Current session info',
    open    = TRUE
  )
```

## References

